---
title: "Learning R3 intermediate"
author: "Sally Søgaard Andersen"
format: html
---

## Setup

```{r setup}
library(tidyverse)
library(here)
library(fs)
source(here("R/functions.R")) # giver os adgang til vores R-script med funktioner
```

## Importing new data

```{r}
#| eval: false #gør at koden ikke kører (beholdes kun til noter ift. læring)

here("data-raw/nurses-stress/stress/5C/5C_1586886626/HR.csv.gz") |>
  read_csv(
    show_col_types = FALSE, # printer ikke dens 'gæt' på kolonnetype
    name_repair = to_snake_case, # gør kolonne-navne til snake_case (nemmere at læse end PascalCase)
    n_max = 100 # bestemmer hvor mange rækker man læser ind fra filen (her 100), fx hvis man gerne vil lave noget prototype-kode hurtigt, så den ikke skal læse et kæmpe dataset ind
  )

# meget kodning hvis man skal skrive dette for hver fil (når der skal indlæses mange filer), så derfor har vi i stedet udarbejdet en funktion (read()), som findes under functions.R
```

## Using map to make hr dataset

```{r}
hr_files <- here("data-raw/nurses-stress/stress/") |>
  dir_ls(regexp = "HR.csv.gz", recurse = TRUE)

hr_data <- hr_files |>
  map(read) |> # ingen (), vi fortæller den at den skal bruge funktionen read, men vi skal ikke selv her aktivt bruge funktionen read()
  list_rbind(names_to = "file_path_id") # rowbind for at samle de mange tibbles til en. Argument names_to giver ID som er lig med filepath
```

## Exercise 18.5 Convert map to function and use on sleep

```{r}
## We made a new function called read_all, can be found under functions.R

hr_data <- read_all("HR.csv.gz")
ibi_data <- read_all("IBI.csv.gz")
bvp_data <- read_all("BVP.csv.gz")
```

## Exercise 19.2 Loading the survey data

```{r}
survey_data <- here("data-raw/nurses-stress/survey-results.csv.gz") |>
  read()
```

## Using regular expressions to extract text

```{r}
hr_data |>
  mutate(
    id = str_extract(
      file_path_id,
      pattern = "/stress/[:alnum:]{2}/" # extracts the complete match
    ) |>
      str_remove("/stress/") |> # fjern /stress/ fra navnet
      str_remove("/"), # fjern / fra navnet
    .before = file_path_id # sæt file_path_id kolonnen først
  ) |>
  select(-file_path_id)
```

## Exercise 19.7: convert get ID code into a function

```{r}
# we created a function called get_participant_id based on the above mentioned regex
# It can be found in functions.R
```

## Exercise 19.8: Simplify the regex with look arounds
```{r}
hr_data |>
  mutate(
    id = str_extract(
      file_path_id,
      pattern = "(?<=/stress/)[:alnum:]{2}(?=/)" # finder det eksakte match
    ),
    .before = file_path_id # sæt file_path_id kolonnen først
  ) |>
  select(-file_path_id)

# har her gjort regex mere præcis og kortere kode. Opdaterer get_participant_id funktionen med denne kode i functions.R
```

## Working with across

```{r}
hr_data <- read_all("HR.csv.gz") |>
  get_participant_id()

hr_data |> summarise(across(where(is.numeric), list(mean = mean, median = median, sd = sd))) # finder numeriske variabler i datasættet og udregner mean, median og sd
```

## Working with dates

```{r}
hr_data |>
  mutate(
    datetime = round_date(collection_datetime, unit = "minute"),
    .before = collection_datetime
  )
```

## Summarising by groups

```{r}
hr_data |>
  mutate(
    collection_datetime = round_date(collection_datetime, unit = "minute")
  ) |>
  summarise(
    across(
      where(is.numeric),
      list(mean = mean, sd = sd, median = median)
    ),
    .by = c(id, collection_datetime)
  ) # split data by and collection_datetime, apply summary statistics and then combine together to output new dataframe.
# .by gør lidt det samme som groupby, det er en nyere funktion
```

## Exercise 20.6 Convert the summarising code into a function

```{r}
# here we created a code to round collecting date time and summarise the dataset. It is in functions.R
```

## Extra exercise 20.8 and 20.9

```{r}
#| eval: false
#' A function to read in any HR-type data
#'
#' @param filename that should be read in and managed
#'
#' @returns summarised dataset with participant id and collection datetime rounded to minutes.
read_sensor_data <- function(filename) {
  data <- read.all(filename) |>
    get_participant_id() |>
    summarise_by_datetime()
  return(data)
}

# you can add max_rows to chance number of rows read
read_sensor_data <- function(filename, max_rows = 100) {
  data <- read_all(filename, max_rows = max_rows) |>
    get_participant_id() |>
    summarise_by_datetime()
  return(data)
}

# you can add units to be able to change rounding
read_sensor <- function(filename, max_rows = 100, unit = "minute") {
  data <- read_all(filename, max_rows = max_rows) |>
    get_participant_id() |>
    summarise_by_datetime(unit = unit)
  return(data)
}
# denne er tilføjet til functions.R
```

## Pivot longer

```{r}
# load survey data
survey_data <- here("data-raw/nurses-stress/survey-results.csv.gz") |>
  read()
survey_data
```

```{r}
# tidying survey-data
survey_tidy <- survey_data |>
  mutate(
    date = mdy(date),
    start_datetime = as_datetime(paste(date, start_time)),
    end_datetime = as_datetime(paste(date, end_time)),
    datetime_id = start_datetime,
    .before = start_time
  ) |>
  select(-c(date, start_time, end_time, duration))
```

```{r}
# pivot longer
survey_tidy |>
  select(id, datetime_id, start_datetime, end_datetime) |>
  pivot_longer(c(start_datetime, end_datetime), names_to = NULL, values_to = "collection_datetime") |>
  group_by(pick(-collection_datetime)) |> # group by every other column besides collection_datetime
  complete(collection_datetime = seq(min(collection_datetime), max(collection_datetime), by = 60)) |>
  ungroup()

# pivot longer og fylder ud så collection_datetime er fyldt med hvert minut, mhp. senere at joine med hr_data
```

## Exercise 21.4

```{r}
# made function to tidy survey data
```

```{r}
# made function to pivot survey data longer
```

## Exercise 21.5 Add survey processing to docs/cleaning.qmd

```{r}
# Findes i cleaning.qmd
```

## Joining survey data to try different joins
```{r}
# Trying different types of data join
survey_long |> left_join(survey_tidy)
survey_long |> right_join(survey_tidy)
survey_long |> full_join(survey_tidy)

# manually inspecting data (finde duplikater)
survey_tidy |>
  count(id, datetime_id) |>
  filter(n > 1)
```
## Join survey data with hr data
```{r}
# Loading hr data in
hr_data <- read_all("HR.csv.gz") |>
  get_participant_id() |>
  summarise_by_datetime()
```

```{r}
nurses_stress <- survey_long |>
  full_join(survey_tidy) |>
  full_join(hr_data)

nurses_stress |>
  arrange(id, collection_datetime) |>
  view()
```

## Exercise 22.5: Join the rest of the data together
```{r}
# in the cleaning doc the dataset have been joined with the ibi data, temp data and bvp data and eda data.
```

